Simulating Dwarf Galaxy Evolution with Individual Star Feedback and Chemical Evolution

### 2018-02-15 main.tex
compiles now

\begin{abstract}
We have developed a new model for stellar feedback in order to simulate the chemodynamical evolution of isolated dwarf galaxies at high resolution in the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. For the first time in galaxy scale simulations, we simulate detailed stellar feedback from individual stars including stellar winds from massive and asymptotic giant branch stars, photoelectric heating, Lyman-Werner dissociation, ionizing radiation tracked through an adaptive ray-tracing radiative transfer method, and core collapse and Type Ia supernovae. 
%This star-by-star method allows us to directly account for variations in stellar properties that govern their feedback properties (mass, lifetime, surface temperature, and metallicity).
We furthermore follow the star-by-star stellar yields using tracer fields for 15 metal species: C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba. We describe these methods in detail, including the additional physics used to model a realistic three-phase interstellar medium. We present the first application of these methods in a 500~Myr simulation of an isolated dwarf galaxy with properties similar to a Local Group ultrafaint dwarf galaxy. We demonstrate that our physics and feedback model is capable of producing a dwarf galaxy whose evolution is consistent with observations in both the Kennicutt-Schmidt relationship and extended Schmidt relationship. Effective feedback drives significantly metal-enriched outflows, leading to low metal retention fractions consistent with observations. Our low-dust content galaxy exhibits a significant H$_2$ mass fraction, formed primarily through gas-phase H$^{-}$ reactions. We end with a discussion on how this work impacts our understanding of galactic chemical evolution on a species-by-species basis. {\bf the conclusion of this discussion should perhaps be mentioned.}
\end{abstract}

@@

\section{Methods}
\label{sec:methods}
%mm2 In  summary, 
We produce high-resolution, galaxy-scale simulations tracking stars not as SSP's, but as individual stars sampled from an assumed IMF.
This allows us to follow star-by-star variations in feedback physics and stellar yields in detail. To properly model the ISM, we track non-equilibrium, primordial chemistry (including molecular hydrogen) using \texttt{Grackle} \citep{GrackleMethod}, with heating and approximate self-shielding from a metagalactic ultraviolet (UV) background. We 
%mm 
    assume collisional ionization equilibrium for all other elements in order to
use updated \texttt{Cloudy} metal-line cooling tables consistent with our self-shielding approximation (see Appendix ~\ref{appendix:cooling}), and an updated observationally motivated dust model for the low metallicity regimes studied here ($Z \lesssim 0.1$ Z$_{\odot}$). Each star is assigned basic properties including surface gravity, effective temperature, radius, and lifetime from tabulated stellar evolution models, which inform how the stars deposit their feedback. We directly track ionizing radiation from massive stars using an adaptive ray tracing radiative transfer method that includes the effects of radiation pressure. In addition, we follow the optically thin, non-ionizing UV radiation from these stars leading to photoelectric heating and Lyman-Werner dissociation of molecular hydrogen. We track the stellar wind feedback and supernovae from these stars, depositing individual metal yields from both. We include AGB wind feedback and yields for lower mass stars, and track these directly as Type Ia supernova progenitors. We follow yields for 15 individual metal species (C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba), chemically tagging each star
%mm
   as it forms
with the associated local gas abundances for each species. In addition, we track a total metal density field which is the sum of all metals, including those not directly tracked. This field is used to inform the heating/cooling physics, and determines the metallicity of each star at birth. These methods are discussed in full detail below.

@@

We use the adaptive mesh refinement (AMR) hydrodynamics and N-body code \texttt{Enzo}\footnote{http://www.enzo-project.org} to simulate the chemodynamical evolution and detailed feedback physics in a set of high resolution, isolated, low-mass dwarf galaxies. \texttt{Enzo} is an open-source code that is undergoing continual, active development by many researchers across several institutions. We use a substantially modified version of the current development version of \texttt{Enzo} (version 2.X) in this work.\footnote{This version is contained in a publicly available fork of the main repository: https://bitbucket.org/aemerick/enzo-emerick. Specifically, simulations presented here were conducted at changeset 445de816fd99.} We solve the equations of hydrodynamics using a direct-Eulerian piecewise parabolic method \citep{ColellaWoodward1984, Bryan1995} and a two-shock approximate Riemann solver with progressive fallback to more diffusive Riemann solvers in the event that higher order methods produce negative densities or energies. We compute the total gravitational potential from gas self-gravity, stars, and a static background dark matter potential (see Section~\ref{sec:IC}). Self-gravity is computed with a multigrid Poisson solver. The collisionless star particles are evolved with an adaptive particle-mesh N-body solver at an effective force resolution of $\sim 2 \Delta x$, where $\Delta x$ is the local cell size.

The computational domain in an AMR simulation is adaptively refined up to a designated resolution limit for a given set of refinement criteria. We require that the Jeans length be resolved by a minimum of 8 cells, continually refining a given region until this criteria is met or the region reaches the maximum resolution. Our maximum resolution in the simulations presented here is 1.8~pc. At maximum resolution, the Jeans length can become under resolved, leading to artificial numerical fragmentation. Studies have shown that resolving the Jeans length by at least 4 cells is required to suppress this fragmentation \citep{Truelove1997}.

@@

@@

\item Formation of a significant amount of H$_2$ through gas phase reactions with H$^_$. This dense gas is more resistive to feedback driven outflows than the less dense gas in the galaxy. This leads to high molecular gas fractions, with an H$_2$ mass comparable to that of \ion{H}{1} by the end of the simulation. This result requires further investigation to determine if this behavior is universally true of ultrafaint dwarf galaxies. See Figure~\ref{fig:ISM_evolution}.
\end{enumerate}

>>>>

\begin{abstract}
We have developed a new model for stellar feedback in order to simulate the chemodynamical evolution of isolated dwarf galaxies at high resolution in the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. For the first time in galaxy scale simulations, we simulate detailed stellar feedback from individual stars including 
%mm2 
     {\bf [should we somehow make clear that we are ignoring energetic feedback from stellar winds?]}
stellar winds from massive and asymptotic giant branch stars, photoelectric heating, Lyman-Werner dissociation, ionizing radiation tracked through an adaptive ray-tracing radiative transfer method, and core collapse and Type Ia supernovae. 
%This star-by-star method allows us to directly account for variations in stellar properties that govern their feedback properties (mass, lifetime, surface temperature, and metallicity).
We furthermore follow the star-by-star stellar yields using tracer fields for 15 metal species: C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba. We describe these methods in detail, including the additional physics used to model a realistic three-phase interstellar medium. We present the first application of these methods in a 500~Myr simulation of an isolated dwarf galaxy with properties similar to a Local Group ultrafaint dwarf galaxy. We demonstrate that our physics and feedback model is capable of producing a dwarf galaxy whose evolution is consistent with observations in both the Kennicutt-Schmidt relationship and extended Schmidt relationship. Effective feedback drives significantly metal-enriched outflows, leading to low metal retention fractions consistent with observations. Our low-dust content galaxy exhibits a significant H$_2$ mass fraction, formed primarily through gas-phase H$^{-}$ reactions. We end with a discussion on how this work impacts our understanding of galactic chemical evolution on a species-by-species basis. {\bf the conclusion of this discussion should perhaps be mentioned.}
\end{abstract}

@@

\section{Methods}
\label{sec:methods}
%mm2 In  summary, 
We produce high-resolution, galaxy-scale simulations tracking stars not as SSP's, but as individual stars sampled from an assumed IMF.
This allows us to follow star-by-star variations in feedback physics and stellar yields in detail. To properly model the ISM, we track non-equilibrium, primordial chemistry (including molecular hydrogen) using \texttt{Grackle} \citep{GrackleMethod}, with heating and approximate self-shielding from a metagalactic ultraviolet (UV) background. We 
assume collisional ionization equilibrium for all other elements in order to
use updated \texttt{Cloudy} metal-line cooling tables consistent with our self-shielding approximation (see Appendix ~\ref{appendix:cooling}), and an updated observationally motivated dust model for the low metallicity regimes studied here ($Z \lesssim 0.1$ Z$_{\odot}$). Each star is assigned basic properties including surface gravity, effective temperature, radius, and lifetime from tabulated stellar evolution models, which inform how the stars deposit their feedback. We directly track ionizing radiation from massive stars using an adaptive ray tracing radiative transfer method that includes the effects of radiation pressure. In addition, we follow the optically thin, non-ionizing UV radiation from these stars leading to photoelectric heating and Lyman-Werner dissociation of molecular hydrogen. We track the stellar wind feedback and supernovae from these stars, depositing individual metal yields from both. We include AGB wind feedback and yields for lower mass stars, and track these directly as Type Ia supernova progenitors. We follow yields for 15 individual metal species (C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba), chemically tagging each star as it forms
with the associated local gas abundances for each species. In addition, we track a total metal density field which is the sum of all metals, including those not directly tracked. This field is used to inform the heating/cooling physics, and determines the metallicity of each star at birth. These methods are discussed in full detail below.

@@

We use the adaptive mesh refinement hydrodynamics and N-body code \texttt{Enzo}\footnote{http://www.enzo-project.org} to simulate the chemodynamical evolution and detailed feedback physics in a set of high resolution, isolated, low-mass dwarf galaxies. \texttt{Enzo} is an open-source code that is undergoing continual, active development by many researchers across several institutions. We use a substantially modified version of the current development version of \texttt{Enzo} (version 2.X) in this work.\footnote{This version is contained in a publicly available fork of the main repository: https://bitbucket.org/aemerick/enzo-emerick. Specifically, simulations presented here were conducted at changeset 445de816fd99.} We solve the equations of hydrodynamics using a direct-Eulerian piecewise parabolic method \citep{ColellaWoodward1984, Bryan1995} and a two-shock approximate Riemann solver with progressive fallback to more diffusive Riemann solvers in the event that higher order methods produce negative densities or energies. We compute the total gravitational potential from gas self-gravity, stars, and a static background dark matter potential (see Section~\ref{sec:IC}). Self-gravity is computed with a multigrid Poisson solver. The collisionless star particles are evolved with an adaptive particle-mesh N-body solver at an effective force resolution of $\sim 2 \Delta x$, where $\Delta x$ is the local cell size.

%mm2 The computational domain in an adaptive mesh refinement simulation is adaptively refined up to a designated resolution limit for a given set of refinement criteria. We require that
     We refine the mesh whenever 
the Jeans length 
%mm2 be
    is no longer
resolved by a minimum of 8 cells, continually refining a given region until this criteria is met or the region reaches the maximum resolution. Our maximum resolution in the simulations presented here is 1.8~pc. At maximum resolution, the Jeans length can become under resolved, leading to artificial numerical fragmentation. \citet{Truelove1997} showed that resolving the Jeans length by at least 4 cells is required to suppress this fragmentation.

@@

%mm2 
   {\bf [I'm a bit confused here: how does the ISRF interact with the cooling?  Shouldn't that move you around in the Cloudy tables, possibly quite substantially?]}

@@

\item Formation of a significant amount of H$_2$ through gas phase reactions with H$^-$. This dense gas is more resistive to feedback driven outflows than the less dense gas in the galaxy. This leads to high molecular gas fractions, with an H$_2$ mass comparable to that of \ion{H}{1} by the end of the simulation. This result requires further investigation to determine if this behavior is universally true of ultrafaint dwarf galaxies. See Figure~\ref{fig:ISM_evolution}.
\end{enumerate}
